MWA tied-array processing III: Microsecond time resolution via a synthesis filter

### 2019-10-22 ipfb.tex
A quick save of progress while I transition to my laptop. Wrote a few more sentences.

The algorithm embodied in the above strategy defines the PFB.
It is clear that choosing a larger value for $P$ will lessen spectral leakage by the same factor, but this naturally comes at a greater computational cost, since the DFT is now applied to $M$ samples.
However, one of the great advantages of the PFB is that the above operations of using a wider analysis wilter and downsampling the resulting spectrum are mathematically equivalent to first segmenting the windowed time series into $P$ segments (known as `taps') of size $N$, summing their respective samples element-wise, and performing a single DFT on the resulting array (now of size $N$), i.e.
\begin{equation}
    X[k] = \sum_{n=0}^{N-1} b[n]\,e^{-2\pi j k (n/N)},
    \label{eqn:pfb}
\end{equation}
where
\begin{equation*}
    b[n] = \sum_{\rho=0}^{P-1} h[n + N\rho]\,x[n + N\rho].
\end{equation*}
A short proof of the equivalence of this expression with the algorithm described above is given in \citet{Harris2011}.
The term `polyphase' derives from the structure of $b[n]$ defined in Eq. \eqref{eqn:pfb}.
It is illustrated in Fig. \ref{fig:pfb}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{example-image-a}
    \caption{A diagramatic representation of the PFB algorithm, as defined in Eq. \eqref{eqn:pfb}.}
    \label{fig:pfb}
\end{figure}

@@

The first stage (coarse) PFB is described in \citet{Prabu2015}, and here we only document a few details pertaining to the second stage (fine) PFB.
A PFB is completely specified by (1) the number of output channels, $N$, (2) the number of taps, $P$, and (3) the analysis filter, $h[m]$, $0 \le m < M=NP$.
For the MWA, $N=128$ (giving fine channels $10\,$kHz wide), $P=12$, and
\begin{equation}
    h[m] = w_H[m]\,w_s[m],
    \label{eqn:filter}
\end{equation}
where
\begin{equation*}
    w_H[m] = \sin^2\bigg(\frac{\pi (m+1)}{M+1}\bigg)
\end{equation*}
is the Hanning window, and
\begin{equation*}
    w_s[m] = \sinc\bigg(\frac{\pi(m + 1 - M/2)}{N}\bigg)
\end{equation*}
is the scaled $\sinc(x) = \sin(x)/x$ function.
It can be easily checked that $h[m]$ is defined to be symmetric around sample $m = 767 = M/2-1$.
The analysis filter is shown in Fig. \ref{fig:filter}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{filter}
    \caption{The coefficients of the analysis filter defined in Eq. \eqref{eqn:filter}, composed of a Hanning window multiplied to a $\sinc$ function.}
    \label{fig:filter}
\end{figure}

>>>>

The algorithm embodied in the above strategy defines the PFB.
It is clear that choosing a larger value for $P$ will lessen spectral leakage by the same factor, but this naturally comes at a greater computational cost, since the DFT is now applied to $M$ samples.
However, one of the great advantages of the PFB is that the above operations of using a wider analysis wilter and downsampling the resulting spectrum are mathematically equivalent to first segmenting the windowed time series into $P$ segments (known as `taps') of size $N$, summing their respective samples element-wise, and performing a single DFT on the resulting array (now of size $N$), i.e.
\begin{equation}
    X[k] = \sum_{n=0}^{N-1} b[n]\,e^{-2\pi j k (n/N)},
    \label{eqn:pfb}
\end{equation}
where
\begin{equation*}
    b[n] = \sum_{\rho=0}^{P-1} h[n + N\rho]\,x[n + N\rho].
\end{equation*}
A short proof of the equivalence of this expression with the algorithm described above is given in \citet{Harris2011}.
The term `polyphase' derives from the structure of $b[n]$ defined in Eq. \eqref{eqn:pfb}.
It is illustrated in Fig. \ref{fig:pfb}.
\begin{figure}[t]
    \centering
    \includegraphics[width=\columnwidth]{example-image-a}
    \caption{A diagramatic representation of the PFB algorithm, as defined in Eq. \eqref{eqn:pfb}.}
    \label{fig:pfb}
\end{figure}

@@

The first stage (coarse) PFB is described in \citet{Prabu2015}, and here we only document a few details pertaining to the second stage (fine) PFB.
A PFB is completely specified by (1) the number of output channels, $N$, (2) the number of taps, $P$, and (3) the analysis filter, $h[m]$, $0 \le m < M=NP$.
For the MWA, $N=128$ (giving fine channels $10\,$kHz wide), $P=12$, and
\begin{equation}
    h[m] = w_H[m]\,w_s[m],
    \label{eqn:filter}
\end{equation}
where
\begin{equation*}
    w_H[m] = \sin^2\bigg(\frac{\pi (m+1)}{M+1}\bigg)
\end{equation*}
is the Hanning window, and
\begin{equation*}
    w_s[m] = \sinc\bigg(\frac{\pi(m + 1 - M/2)}{N}\bigg)
\end{equation*}
is the scaled $\sinc(x) = \sin(x)/x$ function.
It can be easily checked that $h[m]$ is defined to be symmetric around sample $m = 767 = M/2-1$.
The analysis filter is shown in Fig. \ref{fig:filter}.
\begin{figure}[t]
    \centering
    \includegraphics[width=\columnwidth]{filter}
    \caption{Top: The coefficients of the MWA's fine PFB analysis filter, defined in Eq. \eqref{eqn:filter}, which is composed of a Hanning window multiplied to a $\sinc$ function. Bottom: The frequency response of the analysis filter (black, solid), showing negligible attenuation across approximately $10\,$kHz (the bandwidth of a fine channel) and strong attenuation elsewhere. The frequency response is repeated for adjacent channels on either side (grey, dashed) showing crossover points at $-3\,$dB.}
    \label{fig:filter}
\end{figure}

The MWA fine PFB is \textit{critically sampled}, meaning that consecutive $100\,\mu$s output samples are obtained by applying the analysis filter to blocks of samples separated by the length of a tap, i.e. $N$ samples.
This implies that each coarse channel sample becomes part of the input (i.e. a term in the sum) for $P$ consecutive fine channel samples.
However, it should be noted that there is no redundant information.
The term ``critically'' sampled is justified by noting that the downsampling by a factor of \dots inherent in the PFB algorithm is exactly balanced by the
